---
title: "Lab 03 - Nobel laureates"
author: "Sophie Boyd"
date: "1-30-26"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r view-data}
view(nobel)
```

There are 935 observations of 26 variables. Each row represents one Nobel prize. 

### Exercise 2

```{r nobel-living, eval=TRUE}
nobel_copy <- nobel

nobel_living <- nobel_copy %>% filter(
  is.na(died_date),
  !is.na(country),
  gender != "org"
  )

View(nobel_living)
```

Down to 228 observations after filtering for laureates who are still alive, have available information on country, and who are people (not organizations)

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
