---
title: "Lab 03 - Nobel laureates"
author: "Sophie Boyd"
date: "1-30-26"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r view-data}
view(nobel)
```

There are 935 observations of 26 variables. Each row represents one Nobel prize. 

### Exercise 2

```{r nobel-living, eval=TRUE}
nobel_copy <- nobel

nobel_living <- nobel_copy %>% filter(
  is.na(died_date),
  !is.na(country),
  gender != "org"
  )

View(nobel_living)
```

Down to 228 observations after filtering for laureates who are still alive, have available information on country, and who are people (not organizations)

```{r US-other}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r nobel-living-science}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

### Exercise 3

```{r bar-plot}
ggplot(data = nobel_living_science, aes(x = country_us)) +
  geom_bar() +
labs(x = "Location of Scholars When Awarded",
       y = "Number of Nobel Prizes",
     ) +
facet_wrap(~ category, nrow = 4)
```

The plots seem to support the claim in the Buzzfeed article. In each category, there is a higher number of Nobel prizes awarded to scholars in the US than in other countries. The difference is especially stark for prizes in economics.

### Exercise 4

```{r born-US}
nobel_living_science_born_us <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  )

nobel_living_science_born_us %>%
  count(born_country_us, sort = TRUE)
```

105 winners were born in the US.

### Exercise 5

```{r bar-plot-2}
ggplot(data = nobel_living_science_born_us, aes(x = country_us,
                                                fill = born_country_us)) +
  geom_bar() +
labs(x = "Location of Scholars When Awarded",
       y = "Number of Nobel Prizes",
     fill = "Birth Country"
     ) +
facet_wrap(~ category, nrow = 4) 
```

The plots again seem to support the claims in the the Buzzfeed article. Most of the US-based Nobel laureates were born in the US, but there are still sizable groups of Nobel laureates who were awarded in the US but not born in the US.

### Exercise 6

```{r birth-country}
nobel_living_science_born_us %>%
  filter(country == "USA") %>%
  filter(born_country != "USA") %>%
  count(born_country, sort = TRUE)
```

Germany and the United Kingdom are the most common birth countries. 
